Apparent preferences for cognitive effort fade when multiple forms of effort and delay are interleaved in a foraging environment

### 2020-03-20 Cost_manuscript.Rmd
combining result sections

  Both between- and within-subject experiments were pregeristered with the Open Science Framework (https://osf.io/2rsgm/registrations). All experimental procedures were approved by the Boston University Institutional Review Board, and written consent was acquired for all participants. For the between-subject experiment, we recruited individuals until a desired number of 84 eligible participants was achieved (58 Female, median age = 21, range = 18 - 31; number excluded before reaching goal = XX). The sample size was determined by means of power analysis (ANOVA), using a significance level of 0.05, power of 0.8, an effect size of f = 0.45 (calculated from a pilot study), and three groups (one for each cost type). The resulting per-group sample was 20, which we increased to 21 in order to match three possible block orders. We added an extra group of 21 participants who experienced a minimally effortful condition in order to determine whether effort or pure engagement were driving our results.
	We excluded participant datasets based on four preregistered criteria: 1) Consent: If they withdrew their participation; 2) Inattentiveness: a catch trial was placed at the end of each experimental block, asking participants to press a key within 3 seconds (time requirement based on pilot study response times). A participant that failed two or more of these checks was excluded and replaced. 3) Improbable choice behavior: The task was structured so that one reward amount must always be accepted. A participant who quit every trial in at least one block was assumed not to have followed or understood task instructions, or to have disengaged from the task altogether. 4) Performance: In the cognitive effort condition, participants were forced to travel if they made 2 mistakes in a trial (see task procedures below). Any participant with more than 30% forced travels was excluded and replaced.

@@

  Each participant was assigned to one of these types of costs (cognitive effort, physical effort, or delay), and a fourth group of equal size faced an effortless physical task that involved minimal gripping (each group was unaware that other cost conditions existed). On each trial a monetary offer was displayed for 2 s, and participants had the opportunity to expend time and/or effort during the handling time in order to earn it. Upon completion of a trial participants saw a 2 s window displaying the reward obtained, which was followed by a travel time to the next offer. Alternatively, the participant could quit at any point during the handling time by pressing the spacebar on the computer, and immediately start traveling. If participants failed to maintain above-threshold gripping or made two mistakes during the cognitive task, they were forced to travel and missed the reward.
  There were three block types, in which handling times of 2, 10, or 14 seconds were paired with travel times of 14, 6, and 2 seconds, respectively (note that all combinations add up to 16 seconds). Timing parameters were held constant within each 7-min block. Each of these blocks was experienced in a pseudo-randomized order, and repeated in the same order after a short break in order to probe choice stability (total session length = 42 minutes). Reward amounts varied uniformly per trial (4, 8, or 20 cents), such that every reward was presented twice every six trials. This prevented sequences from being dominated by a single amount during any window of time. Timing information was disclosed at the beginning of each block, and rewards displayed before each trial began during a 2 s window. Participants received training prior to the experimental session, and were told about all possible environmental statistics in order to preclude experience-dependent learning. 

@@

  First, we tested whether decision makers integrated delay and reward information. To address the prediction that participants would be more likely to accept higher reward and shorter handling time trials, we fit a logistic regression to each participant’s data to predict trial-wise acceptances, using handling time and reward amount as predictors. The resulting beta coefficients for handling time and reward were pooled across all participants, and we performed a one-sample rank-sum test on each set of coefficients to examine whether they were significantly positive or negative (compared to no effect of 0). We then performed an extension of the logistic regression by adding an autoregressive covariate containing the number of consecutive quits prior to a given trial. This examined the possibility that participant choices were governed by recent quitting history rather than the experimental parameters. Coefficients significantly different from 0 denoted that a participant relied on recent quitting history. 

@@

  Once we established choice stability, we proceeded to estimate the subjective OC of each type of demand (see Operationalization of Cost above). OC were computed as the product of a free parameter (gamma) and the handling time. We used a softmax function to model each participant’s probability of completing a trial based on the difference between the delayed reward’s magnitude and the estimated OC for each cost type.

@@

### BTW: 16.1.1.

  *To address hypothesis 4.1., A logistic regression will be fit for each participant in order to predict trial-wise acceptances, using handling time and reward amount as predictors. The resulting beta coefficients for handling time and reward will be pooled across all participants, and we will perform a one-sample rank-sum test on each set of coefficients to examine whether the they are significantly positive or negative (compared to 0). If the group coefficients are significantly positive, it would mean that a predictor reliably increases the likelihood of acceptance. This will allow us to determine whether increments in handling time and reward amounts increased and decreased the likelihood of acceptance for each participant, respectively.*

``` {r BTW: 16.1.1., fig.align="center", fig.width = 4, fig.height = 4, echo = FALSE}
# filterout errors and compute R + H logistic per subject, then get coefficients
logisRH <- dataBtw %>%
  filter(Choice < 2) %>%
  plyr::dlply("SubjID", identity) %>%
  lapply(function(data) {glm(Choice ~ Handling + Offer, data = data, family = "binomial")}) %>%
  sapply(coefficients)

@@

The plot shows the pooled coefficients across subjects for the effect of each predictor. The results show the expected discounting effect of increasing the handling time, as well as the increase in acceptance likelihood as a function of reward increments. We found that these coefficients were significantly different from zero for both handling (V = `r rankHand$statistic`, p `r ifelse(rankHand$p.value < 0.001, "< 0.001", paste("=", round(rankHand$p.value, digits = 2)))`) and reward (V = `r rankOffer$statistic`, p `r ifelse(rankOffer$p.value < 0.001, "< 0.001", paste("=", round(rankHand$p.value, digits = 2)))`) regardless of cost condition.

### BTW: 16.1.2.

*We will perform an extension of the logistic regression from 16.1.1., this time adding an autoregressive covariate containing the number of consecutive quits prior to a given trial. In this way, we will examine the possibility that participant choices were governed by recent quitting history rather than the experimental parameters (see 11.1.3.). Coefficients not significantly different from 0 will denote that a participant did not rely on recent quitting history.*

``` {r BTW: 16.1.2., echo = FALSE, include = FALSE}
# add an AR regressor to track recent quits
logisRHAR <- dataBtw %>%
              filter(Choice < 2) %>%
              group_by(SubjID) %>%
              mutate(AR = seqQuits(Choice)) %>%
              plyr::dlply("SubjID", identity) %>%
              lapply(function(data) {glm(Choice ~ Handling + Offer + AR, data = data, family = "binomial")})

# Rank sums
rankAR <- logisRHAR %>% 
  sapply(coefficients) %>%
  wilcox.test(.["AR", ])

AR_CIs <- sapply(logisRHAR, function(x) {tryCatch(confint(x)[4, ], error = function(e){c(0, 0)})}) %>%
  replace_na(0) %>%
  t() %>% 
  apply(1, function(x) !(x[1] <= 0 & x[2] >= 0))

@@

To measure the significance of each subject's autoregressive coefficient, we computed its CI and checked how many contained zero. By this measure, `r sum(AR_CIs)` out of `r nSubjs_btw` of our participants seemed to have been influenced by recent quitting. However, the autoregressive predictor did not preclude the effect of the remaininig experimental parameters.

@@

>>>>

  Both between- and within-subject experiments were pregeristered with the Open Science Framework (https://osf.io/2rsgm/registrations). All experimental procedures were approved by the Boston University Institutional Review Board, and written consent was acquired for all participants. For the between-subject experiment, we recruited individuals until a desired number of 84 eligible participants was achieved (58 Female, median age = 21, range = 18 - 31; number excluded before reaching goal = XX). The sample size was determined by means of power analysis (ANOVA), using a significance level of 0.05, power of 0.8, an effect size of f = 0.45 (calculated from a pilot study), and three groups (one for each cost type). The resulting per-group sample was 20, which we increased to 21 in order to match three possible block orders. We added an extra group of 21 participants who experienced a minimally effortful condition in order to determine whether effort or pure engagement were driving our results.
	We excluded participant datasets based on four preregistered criteria: 1) Consent: If they withdrew their participation; 2) Inattentiveness: a catch trial was placed at the end of each experimental block, asking participants to press a key within 3 seconds (time requirement based on pilot study response times). A participant that failed two or more of these checks was excluded and replaced. 3) Improbable choice behavior: The task was structured so that one reward amount must always be accepted. A participant who quit every trial in at least one block was assumed not to have followed or understood task instructions, or to have disengaged from the task altogether. 4) Performance: In the cognitive effort condition, participants were forced to travel if they made 2 mistakes in a trial (see task procedures below). Any participant with more than 30% forced travels was excluded.

@@

  Each participant was assigned to one of these types of costs (cognitive effort, physical effort, or delay), and a fourth group of equal size faced an effortless physical task that involved minimal gripping. Each group was unaware that other cost conditions existed. On each trial a monetary offer was displayed for 2 s, and participants had the opportunity to expend time and/or effort during the handling time in order to earn it. Upon completion of a trial participants saw a 2 s window displaying the reward obtained, which was followed by a travel time to the next offer. Alternatively, the participant could quit at any point during the handling time by pressing the spacebar on the computer, and immediately start traveling. If participants failed to maintain above-threshold gripping or made two mistakes during the cognitive task, they were forced to travel and missed the reward.
  There were three block types, in which handling times of 2, 10, or 14 seconds were paired with travel times of 14, 6, and 2 seconds, respectively (note that all combinations add up to 16 seconds). Timing parameters were held constant within each 7-min block. Each of these blocks was experienced in a pseudo-randomized order, and repeated in the same order after a short break in order to probe choice stability (total session length = 42 minutes). Reward amounts varied uniformly per trial (4, 8, or 20 cents), such that every reward was presented twice every six trials. This prevented sequences from being dominated by a single amount during any window of time. Timing information was disclosed at the beginning of each block, and rewards displayed during a 2 s offer window before each trial began. Participants received training prior to the experimental session, and were told about all possible environmental statistics in order to preclude experience-dependent learning. 

@@

  First we tested whether decision makers integrated delay and reward information. To address the prediction that participants would be more likely to accept higher reward and shorter handling time trials, we fit a logistic regression to each participant’s data to predict trial-wise acceptances, using handling time and reward amount as predictors. The resulting beta coefficients for handling time and reward were pooled across all participants, and we performed a one-sample rank-sum test on each set of coefficients to examine whether they were significantly positive or negative (compared to no effect of 0). We then performed an extension of the logistic regression by adding an autoregressive covariate containing the number of consecutive quits prior to a given trial. This examined the possibility that participant choices were governed by recent quitting history rather than the experimental parameters. Coefficients significantly different from 0 denoted that a participant relied on recent quitting history. 

@@

  Once we established choice stability, we proceeded to estimate the subjective OC of each type of demand (see Operationalization of Cost above). OC were computed as the product of a free parameter (gamma) and the handling time. We used a softmax function to model each participant’s probability of completing a trial based on the difference between the delayed reward’s magnitude and the estimated OC for each cost type

@@

``` {r BTW: 16.1.1., fig.align="center", fig.width = 4, fig.height = 4, echo = FALSE}
# filter out errors and compute R + H logistic per subject, then get coefficients
logisRH <- dataBtw %>%
  filter(Choice < 2) %>%
  plyr::dlply("SubjID", identity) %>%
  lapply(function(data) {glm(Choice ~ Handling + Offer, data = data, family = "binomial")}) %>%
  sapply(coefficients)

# coefficient summaries
coeffRH_mean <- apply(logisRH, 1, mean)
coeffRH_se <- apply(logisRH, 1, sd) / sqrt(nSubjs_btw)

@@

``` {r BTW: 16.1.2., echo = FALSE, include = FALSE}
# add an AR regressor to track recent quits
logisRHAR <- dataBtw %>%
              filter(Choice < 2) %>%
              group_by(SubjID) %>%
              mutate(AR = seqQuits(Choice)) %>%
              plyr::dlply("SubjID", identity) %>%
              lapply(function(data) {suppressWarnings(glm(Choice ~ Handling + Offer + AR, data = data, family = "binomial"))})

# get the AR CIs per participant 
AR_CIs <- sapply(logisRHAR, function(x) {tryCatch(suppressWarnings(confint(x)[4, ]), error = function(e){c(0, 0)})}) %>%
  replace_na(0) %>%
  t() %>%
  apply(1, function(x) (x[1] <= 0 & x[2] >= 0))

# coefficient summaries
logisRHAR <- sapply(logisRHAR, coefficients)
coeffRHAR_mean <- apply(logisRHAR, 1, mean, na.rm = T)
coeffRHAR_se <- apply(logisRHAR, 1, sd, na.rm = T) / sqrt(nSubjs_btw)

# Rank sum test
rankRHAR <- wilcox.test(logisRHAR["AR", ])

@@

  Before comparing acceptance rates among cost conditions, we determined whether participants integrated the rewarding statistics of the experimental environment. We first confirmed that increasing handling times and reward amounts produced the expected reduction and increase in acceptance probabilities, respectively. Logistic regression coefficients for handling time were significantly lower than 0 across participants (mean $\beta$ = `r mean(rankAR, na.rm = T)`, SE = `r coeffRH_se[2]`; Wilcoxon test: V = `r rankHand$statistic`, p `r ifelse(rankHand$p.value < 0.001, "< 0.001", paste("=", round(rankHand$p.value, digits = 2)))`), whereas increasing the reward amount resulted in higher probabilities (mean $\beta$ = `r coeffRH_mean[1]`, SE = `r coeffRH_se[1]`; Wilcoxon test: V = `r rankOffer$statistic`, p `r ifelse(rankOffer$p.value < 0.001, "< 0.001", paste("=", round(rankHand$p.value, digits = 2)))`). We next examined whether choices could also have been affected by recent choice history (i.e. if too many sequential quits promoted acceptances regardless of reward amount) by adding a regressor to the model that tracked the number of quits preceding a trial. This auto-regressive coefficient was not significantly different than 0 across the whole group (mean $\beta$ = `r coeffRHAR_mean[3]`, SE = `r coeffRHAR_se[3]`; Wilcoxon test: V = `r rankRHAR$statistic`, p `r ifelse(rankRHAR$p.value < 0.001, "< 0.001", paste("=", round(rankRHAR$p.value, digits = 2)))`). Further, the 95% confidence interval (CI) of these coefficients contained 0 in `r round((sum(AR_CIs) / nSubjs_btw) * 100)`% of our participants.

@@

CONCLUDING: these results suggest that participants were adequately influenced by the rewarding statistics of the environment, even though they did not always perform optimally.
